---
title: "Lab 3: COVID-19"
subtitle: 'Ecosystem Science and Sustainability 330'
author:
  - name: Neva Morgan
    email: neva.morgan@colostate.edu
format: html
---

```{r}
library(tidyverse)
library(zoo)
library(flextable)

url <- 'https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'
covid <- read.csv(url)

```

### **Question 1:** **Public Data**

#### Take a moment to reflect on the value of open data:

1.  How does easy access to historical and real-time environmental, data shape our understanding of climate trends, resource management, and public health?
    1.  It's crucial for data especially concerning climate trends, resource management, and public health to be public information, but I can understand the concern of data privacy as it relates to foreign body intentions. Certain aspects of data are vital to be kept private, for example, even data that is pulled from a CitSci organization must have a check of intentions with the data they are requesting. I think for future projections of any real data that is interacting with people internationally, that should be public knowledge. Giving the ability of access is crucial in developing possible new strategies for combating and mitigating lasting effects of climate change.
2.  What happens when this data disappears or becomes inaccessible?
    1.  
3.  The role of independent archiving and collaborative stewardship has never been more critical in ensuring scientific progress and accountability.
    1.  

\~\~\~

### **Question 2: Daily Summary**

Focused criteria to use to place countries on the watch list for COVID-19:

-   More than 100 new cases per 100,000 residents over the past 14 days…

OBJECTIVES:

1.  cumulative cases in the 5 worst counties

    1.  sum?

2.  total **NEW** cases in the 5 worst counties

    1.  max(date)

3.  A list of safe counties

    1.  any country with less than 100 new cases per 100,000 residents over the past 14 days.

4.  A text report describing the total new cases, total cumulative cases, and number of safe counties.

```{r}

```

# **Question 6: Multi-state**

## **Steps:**

```{r}

```

1.  First, we need to `group/summarize` our county level data to the state level, `filter` it to the four states of interest, and calculate the number of daily new cases (`diff/lag`) and the 7-day rolling mean.

**Rolling Averages**

**Hint:** You will need two `group_by` calls and the `zoo::rollmean` function.

2.  Using the modified data, make a facet plot of the daily new cases and the 7-day rolling mean. Your plot should use compelling geoms, labels, colors, and themes.

<!-- -->

3.  The story of raw case counts can be misleading. To understand why, lets explore the cases per capita of each state. To do this, join the state COVID data to the population estimates and calculate the newcases/totalpopulation. Additionally, calculate the 7-day rolling mean of the new cases per capita counts. **This is a tricky task and will take some thought, time, and modification to existing code (most likely)!**

**Hint**: You may need to modify the columns you kept in your original population data. Be creative with how you join data (inner vs outer vs full)!

4.  Using the per capita data, plot the 7-day rolling averages overlying each other (one plot) with compelling labels, colors, and theme.

<!-- -->

5.  Briefly describe the influence scaling by population had on the analysis? Does it make some states look better? Some worse? How so?
